---
#title: "Carwash Analytics"
#author: "Tshepiso M"
#date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}

#This chunk runs first and sets up everything
knitr::opts_chunk$set(
  echo = FALSE,      # Don't show code in final report
  warning = FALSE,   # Hide warnings
  message = FALSE    # Hide messages
)
# Load required libraries
library(dplyr)       # Provides the %>% pipe operator
library(knitr)       # Provides the kable function
library(kableExtra)  # Provides kable_styling, column_spec, etc.
```

<!-- ====== BANNER ====== -->

<div style="background: linear-gradient(135deg, #1a5276 0%, #2e86c1 100%); 
            color: white; 
            padding: 30px 40px; 
            border-radius: 12px; 
            margin-bottom: 30px;
            text-align: center;
            box-shadow: 0 4px 15px rgba(0,0,0,0.2);">

  <h1 style="margin: 0; font-size: 42px; font-weight: 700;"> Statistical Analysis Report </h1>
  <p style="margin: 8px 0 0 0; font-size: 18px; opacity: 0.9;">
    Statistics for Operational Excellence
  </p>
  <hr style="width: 100px; border: 1px solid rgba(255,255,255,0.3); margin: 15px auto;">
  <p style="margin: 0; font-size: 14px; opacity: 0.8;">
    Prepared by: Tshepiso M &nbsp;|&nbsp; Date: `r format(Sys.Date(), "%B %d, %Y")`
  </p>
</div>

<!-- ====== EXECUTIVE SUMMARY ====== -->
<div style="background: #f8f9fa; 
            padding: 20px 25px; 
            border-radius: 10px; 
            border-left: 5px solid #2e86c1;
            margin-bottom: 30px;
            font-size: 16px;
            line-height: 1.6;">


This report provides a comprehensive analysis of a carwash operational performance enabling strategic improvements and growth opportunities.

</div>


<!-- ====== Statistics summary ====== -->

<style>

  /* Applied to ALL tabs (sets fixed width and alignment) */
  .nav-pills > li > a {
    width: 240px !important;       /* Adjust width as needed */
    text-align: center !important;  /* Centers text inside fixed box */
    border-radius: 6px;
    font-weight: 600;
  }
</style>

## 1. Customer Statistical Analysis
### {.tabset .tabset-pills}

#### Lifetime Value Summary Table

```{r}

#Read data
setwd('C:/Users/tshep/mobile_carwash/data/statistics/customer_metrics')
ltv_summary=read.csv('ltv_summary.csv',header = TRUE)

ltv_summary %>%
  kable(
    caption = "Customer Lifetime Value  Statistical Summary",
    digits = 2
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = TRUE)
```

#### Insights

<!-- ====== Mean evaluation ====== -->

<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-width: 4px;">
  <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 600; font-size: 15px;"> Average Lifetime Value</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
    The average customer lifetime value sits at <strong>1,191.02</strong>, exceeding the median of <strong>880.00</strong>.
This tells us that a smaller group of "high value" customers is pulling the average up, while the majority of your customer base generates a lower lifetime value. 

This is further proven by standard deviation value of <strong>1054.11</strong>, which informs us that there's an extreme inequality in customer lifetime values. We can then formulate a hypothesis that our data is right skewed until proven otherwise.

The mean value of <strong>1,191.02</strong> cannot be reliable. We can instead opt for the median to make operational decisions 
  </p>
</div>


<!-- ====== shape distribution ====== -->

<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-width: 4px;">
  <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 600; font-size: 15px;"> Shape of Distribution</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
The skewness value resulted to <strong>1.13</strong> while the kurtosis value sits at <strong>4.04</strong>.

This skewness value is over 2 times our benchmark of <strong>0.5</strong>. The statistic proves our hypothesis that our data is positively skewed.
On the other hand, the kurtosis value of <strong>4.04</strong> shows the presence of a heavy tail on the right which means extreme values do exist.
  </p>
</div>




### {.tabset .tabset-pills}

#### Confidence Intervals

```{r}

#Read data
setwd('C:/Users/tshep/mobile_carwash/data/statistics/customer_metrics')
mean_ltv_ci=read.csv('mean_ltv_ci.csv',header = TRUE)

mean_ltv_ci %>%
  kable(
    caption = "Customer Lifetime Value confidence interval",
    digits = 2
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = TRUE)
```

```{r}

#Read data
setwd('C:/Users/tshep/mobile_carwash/data/statistics/customer_metrics')
repeat_rate_ci=read.csv('repeat_rate_ci.csv',header = TRUE)

repeat_rate_ci %>%
  kable(
    caption = "Booking Repeat Rate confidence interval",
    digits = 2
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = TRUE)


```


#### Insights
<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-width: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">1. Lifetime Value Confidence Interval</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
  
The shape of the distribution informs us that the data does not follow a normal distribution since it is highly skewed, however, the number of data is greater than 30, <strong>n>30</strong>. This triggers 2 theorems:

<strong>1. Central Limit Theorem:</strong> Sampling distribution of the mean becomes approximately normal, regardless of the original population.

<strong>2. Slutsky's Theorem:</strong> Sample standard deviation becomes a very accurate, stable estimate of the population standard deviation,σ.

Therefore t-test becomes reliable for our confidence interval calculations

Even the absolute worst-case estimate for the population mean <strong>888.24</strong> is higher than what the typical customer actually spends, <strong>880.00</strong>.

Our typical chosen average lifetime value of <strong>880.00</strong>[median] does not fall within the the calculated confidence interval. This proves that taking "confidence" in the mean is dangerous.

Even at the lowest bound of our confidence interval, we would still be overestimating the lifetime value generated by the majority of customers.

  </p>
</div>






<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-widt h: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">2. Repeat Rate Confidence Interval</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
  
Calculated Repeat Rate amounted to <strong>93.9%</strong>. Nearly 94% of customers return for more than one booking.
We are therefore 95% Confident that the true Repeat Rate in the interval <strong>[82.1%, 98.4%]</strong>.

Upper bound is close to 100%. We can therefore conclude that <strong>Customer Retention</strong>.Is almost perfect.

  </p>
</div>

## 2. Revenue Statistical Analysis

### {.tabset .tabset-pills}

#### Daily Revenue Trend Analysis
```{r}

#REQUIRED PACKAGES
library(tidyverse)
library(ggplot2)

#READ DATA
setwd('C:/Users/tshep/mobile_carwash/data/statistics/revenue_metrics')
daily_revenue=read.csv("daily_revenue.csv", header = TRUE)
zero_period=read.csv("zero_period.csv", header = TRUE)

#DATE CONVERSIONS
daily_revenue$payment_date=as.Date(daily_revenue$payment_date)
zero_period$payment_date=as.Date(zero_period$payment_date)

closure_start= min(zero_period$payment_date)
closure_end =  max(zero_period$payment_date)
closure_mid = mean(c(closure_start, closure_end))

#DAILY TREND PLOTTING
ggplot(daily_revenue, aes(x=payment_date, y=payment_amount))+
  geom_line(color="steelblue", linewidth=0.6)+
  annotate("rect",
           xmin = closure_start,  
           xmax = closure_end,
           ymin = 0,
           ymax = Inf,
           alpha=0.2,
           fill="red")+
  annotate("text",
           x = closure_mid,
           y = max(daily_revenue$payment_amount) * 0.8,
           label = "No Sales (Closure/Maintenance)",
           color="red",
           fontface="bold")+
           labs(
              title = "Daily Revenue Trend ",
              x = "Date",
              y = "Revenue (R)") +theme_minimal()

```


#### Insights


<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-widt h: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">2-Month Revenue Gap Analysis</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">

The business experienced a significant 2-Month Gap in revenue from April to mid June. This period reflects operational downtime due to maintenance and renovations. 


  </p>
</div>
### -
### -
### {.tabset .tabset-pills}

#### Revenue Summary Table
```{r}

#Read data
setwd('C:/Users/tshep/mobile_carwash/data/statistics/revenue_metrics')
revenue_summary=read.csv('revenue_summary.csv',header = TRUE)

revenue_summary %>%
  kable(
    caption = "Revenue Statistical Summary: All Data",
    digits = 2
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = TRUE)
```

<!-- ====== PROCESS OF REMOVING ZERO PERIOD GAP(MAINTENANCE) FROM THE MAIN DATASET ====== -->
<div style="background: #f8f9fa; 
            padding: 20px 25px; 
            border-radius: 10px; 
            border-left: 5px solid #2e86c1;
            margin-bottom: 30px;
            font-size: 16px;
            line-height: 1.6;">


The two-month revenue gap (April–May 2026) was excluded to prevent distortion of average daily revenue and trend analysis. This ensures that all metrics reflect only active trading periods, allowing for accurate performance assessment and a clear quantification of revenue lost during the downtime.

</div>
### -
### -
### Active Period Analysis
### {.tabset .tabset-pills}

#### Daily Revenue Trend Analysis (Active Period)
```{r}

#REQUIRED PACKAGES
library(tidyverse)
library(ggplot2)

#READ DATA
setwd('C:/Users/tshep/mobile_carwash/data/statistics/revenue_metrics')
daily_revenue=read.csv("daily_revenue.csv", header = TRUE)
zero_period=read.csv("zero_period.csv", header = TRUE)

gap_start=min(zero_period$payment_date)
gap_end=max(zero_period$payment_date)
revenue_daily_active=daily_revenue[daily_revenue$payment_date < gap_start |
                                     daily_revenue$payment_date > gap_end,]


#CONVERT DATES TO FACTOR: TO OFFICIALLY REMOVE THE GAP IN THE PLOTS
revenue_daily_active$date_factor=as.factor(revenue_daily_active$payment_date)

ggplot(revenue_daily_active, aes(x = date_factor, y = payment_amount, group=1)) +
  geom_line(color = "steelblue", linewidth = 0.6) +
  labs(
    title = "Daily Revenue Trend (Active Period)",
    x = "Date",
    y = "Revenue (R)"
  )+theme_minimal() +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 15)])
```


#### Histogram (Active Period)
```{r}

#REQUIRED PACKAGES
library(tidyverse)
library(ggplot2)

#READ DATA
setwd('C:/Users/tshep/mobile_carwash/data/statistics/revenue_metrics')
daily_revenue=read.csv("daily_revenue.csv", header = TRUE)
zero_period=read.csv("zero_period.csv", header = TRUE)

gap_start=min(zero_period$payment_date)
gap_end=max(zero_period$payment_date)
revenue_daily_active=daily_revenue[daily_revenue$payment_date < gap_start |
                                     daily_revenue$payment_date > gap_end,]


ggplot(revenue_daily_active, aes(x=payment_amount))+
  geom_histogram(binwidth = 200, boundary=0,  fill="lightblue", color="white")+
  #DENSITY LINE
  geom_density (aes(y=after_stat(density)*nrow(revenue_daily_active)*200),
                
                color="firebrick", 
                linewidth=1)+
  labs(title="Distribution of Daily Revenue (Active Periods)",
       x="Revenue",
       y="Density")+
  theme_minimal()
```


#### Revenue Summary Table (Active period)
```{r}

#Read data
setwd('C:/Users/tshep/mobile_carwash/data/statistics/revenue_metrics')
active_revenue_summary=read.csv('active_revenue_summary.csv',header = TRUE)

active_revenue_summary %>%
  kable(
    caption = "Revenue Statistical Summary: Active Period",
    digits = 2
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = TRUE)
```

#### Insights


<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-widt h: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">Comparison Table: Active data vs Active Period</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">

Metric              |All Data   |Active Period |Diferrence
--------------------|-----------|--------------|--------------
Total Revenue       | R58 360    |R58 360       |R0
Mean Daily Revenue  | R284.68    |R459.53       |+R174.85
Median Daily Revenue| R0         |R440.00       |+R440.00
Standard Deviation  | R409.23    |R436.07       |+R26.84
Skewness            | 1.30       |0.64          |	-0.66
Kurtosis            | 3.72       |2.55          |-1.17
Min                 | R0         |R0            | -
Max                 | R1 640     |R1 640        | -

  </p>
</div>

<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-widt h: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">Remarks</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
*  When excluding the two-month revenue gap (April–May 2026), the average daily revenue increases from <strong>R284.68</strong>. to <strong>R459.53 </strong>, 61% increase.

*  The distribution also becomes less skewed (skewness <strong>0.64</strong> vs <strong>1.30</strong>)

* And less prone to extreme outliers (kurtosis <strong>2.55</strong> vs <strong>3.72</strong>).

*  This confirms that the gap significantly distorted the overall revenue picture, and active-period metrics provide a more accurate baseline for operational performance and loss estimation.

* Statistical tests needs to be conducted in order to verify our confidence in our mean for the active period

  </p>
</div>




<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-widt h: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">Loss Calculation</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
  

  </p>
</div>

### -
### Normality Assessment (Active Period)

### {.tabset .tabset-pills}
#### Q-Q Plot
```{r}

#Read data
setwd('C:/Users/tshep/mobile_carwash/data/statistics/revenue_metrics')
revenue_daily_active=read.csv('revenue_daily_active.csv',header = TRUE)

qqnorm(revenue_daily_active$payment_amount,
       main = "Normal Q-Q Plot: Daily Revenue (Active Period) ",
       xlab = "Theorical Quantiles",
       ylab = "Sample Quantiles")
qqline(revenue_daily_active$payment_amount, col="red", lwd=2)
```

#### Shapiro-Wilk Test
```{r}

#Read data
setwd('C:/Users/tshep/mobile_carwash/data/statistics/revenue_metrics')
shapiro_test=read.csv('shapiro_test.csv',header = TRUE)


    
shapiro_test %>%
    mutate(
    p_value = ifelse(p_value < 0.001, "< 0.001", round(p_value, 3))
  ) %>%
  kable(
    caption = "Shapiro-Wilk Test",
    digits = 2
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = TRUE)
```

#### Insights

<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-widt h: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">Normality Remark: Q-Q Plot</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
*  Portion of theorical quantiles, <strong>-2.8 < X < -0.4</strong> lies flat at a sample quantile of 0. This indicates a high concentration of zero-revenue days

*  While the middle-to-upper values roughly follow the reference line, the far right end curves slightly below the line, indicating a light upper tail relative to a normal distribution.
  </p>
</div>


<div style="background: #ffffff; border-left: 4px solid #1a5276; padding: 16px 20px; margin-top: 20px; border-radius: 0 8px 8px 0; box-shadow: 0 2px 8px rgba(0,0,0,0.04); border: 1px solid #eef2f6; border-left-widt h: 4px;">

 <h5 style="color: #1a5276; margin: 0 0 8px 0; font-weight: 630; font-size: 15px;">Sapiro-Wilk Test</h5>
  <p style="margin: 0; color: #4a5568; font-size: 13.5px; line-height: 1.6;">
*   <strong>P-Value < 0.05</strong>. Hypothesis for Normality is rejected, and we can therefore conclude that our data does not follow normality.

*  However, our data, <strong> n > 100</strong>, the Central Limit Theorem Applies. We can conclude that our sample statistic, mean approximates a normal distribution regardless of the intial distribution. Parametric Tests are eligible.

  </p>
</div>
